\normalfont\spacedallcaps{Language reclamation: Literature review}

### 2023-06-18 assignment-4/article.tex
Solve 4.3
Add the questions for the rest to be solved

\subsection*{Exercise 4.4 (Normalizing Flows: Fixed point theorem)}
Problem 16.11 from Chapter 16 of ”Understanding Deep Learning”

\subsection*{Exercise 4.5 (Diffusion generative models)}
Problems 18.1 and 18.2 from Chapter 18 of ”Understanding Deep Learning”

>>>>

\subsubsection*{Problem 16.2}

Consider transforming a standard normal distribution:

\begin{equation*}
  Pr(z) = \frac{1}{\sqrt{2\pi}} \exp \left[\frac{-z^2}{2}\right]
\end{equation*}

with the function:

\begin{equation*}
  x = f[z] = \frac{1}{1 + \exp[-z]}
\end{equation*}

Find an expression for the transformed distribution $Pr(x)$.

\subsubsection*{Problem 16.2 solution}

As before, let $P_Z = Pr(z)$. Then $x = f[z]$ implies $z = \log(x(1-x))$, from which we derive: $\frac{dz}{dx} = \frac{1-2x}{x(1-x)}$

Then using the change of variables formula from before:

\begin{align*}
  p_X(x) &= p_Z(z) \left|\frac{dz}{dx}\right| \\
  &= \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{1}{2}(\log(x(1-x)))^2 \right] \cdot \frac{1-2x}{x(1-x)} \\
  &= \frac{1}{\sqrt{2\pi}} \exp \left[\log(x(1-x)))^{-\frac{1}{2} \cdot 2} \right] \cdot \frac{1-2x}{x(1-x)} \\
  &= \frac{1}{\sqrt{2\pi}} \exp \left[\log\left(\frac{1}{x(1-x)}\right) \right] \cdot \frac{1-2x}{x(1-x)} \\
  &= \frac{1}{\sqrt{2\pi}} \frac{1}{x(1-x)} \cdot \frac{1-2x}{x(1-x)} \\
  &= \frac{1}{\sqrt{2\pi}} \frac{1-2x}{x^2(1-x)^2}
\end{align*}

%% \subsubsection*{Problem 16.2 solution - attempt 2}

%% Given the standard normal distribution:
%% \begin{equation*}
%%   Pr(z) = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{z^2}{2}\right]
%% \end{equation*}

%% And the function:
%% \begin{equation*}
%%   x = f[z] = \frac{1}{1 + \exp[-z]}
%% \end{equation*}

%% We can derive the inverse function $f^{-1}(x)$:
%% \begin{equation*}
%%   z = f^{-1}(x) = -\log \left(\frac{1}{x} - 1\right)
%% \end{equation*}

%% The formula for the change of variables in probability densities is:

%% \begin{equation*}
%%   Pr(x) = Pr(f^{-1}(x)) \left| \frac{df^{-1}(x)}{dx} \right|
%% \end{equation*}

%% By differentiating $f^{-1}(x)$ with respect to $x$, we get $\frac{df^{-1}(x)}{dx} = \frac{1}{x(1-x)}$.

%% Substituting $f^{-1}(x)$ and $\frac{df^{-1}(x)}{dx}$ into the formula, we get:

%% \begin{equation*}
%%   Pr(x) = \frac{1}{\sqrt{2\pi}} \exp \left[-\frac{(-\ln(\frac{1}{x} - 1))^2}{2}\right] \cdot \frac{1}{x(1-x)}
%% \end{equation*}

%% This is the probability density function of the transformed distribution.

\subsection*{Exercise 4.4 (Normalizing Flows: Fixed point theorem)}

\subsubsection*{Problem 16.11}

Write out the expression for the KL divergence in the equation:

\begin{equation*}
  \phi = \argmin \left[ KL \left[ \sum_{i=1}^I \delta[x-f[z_i,\phi]] || q(x) \right] \right]
\end{equation*}

Why does it not matter if we can only evaluate the probability q(x) up to a scaling factor $\kappa$? Does the network have to be invertible to minimize this loss function? Explain your reasoning.

\subsection*{Exercise 4.5 (Diffusion generative models)}

\subsubsection*{Problem 18.1}

Show that if $Var[x_{t-1}] = \textbf{I}$ and we use the update:

\begin{equation*}
  x_t = \sqrt{1-\beta_t}\cdot x_{t-1} + \sqrt{\beta_t} \cdot \epsilon_t,
\end{equation*}

then $Var[x_t] = \textbf{I}$ so the variance stays the same.

\subsubsection*{Problem 18.2}

Consider the variable:

$$z = a \cdot \epsilon_1 + b \cdot \epsilon_2$$

where both $\epsilon_1$ and $\epsilon_2$ are drawn from independent standard normal distributions with mean zero and unit variance. Show that:

\begin{align*}
  \mathbb{E}[z] &= 0 \\
  Var[z] &= a^2 + b^2
\end{align*}

so we could equivalently compute $z = \sqrt{a^2 + b^2} \cdot \epsilon$ where $\epsilon$ is also drawn from a standard normal distribution.
